Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey

### 2014-01-21 mard-survey.tex
2012.05.29 (Vassilis) Monolithic file broken down to multiple smaller ones (per section) for better parallel contribution. Additional material on more recent algorithms for unstructured networks. Now also covering the BitTorrent protocol. Merged Huseyin's figures from previous contribution and code refactored for better referencing. Fixed bibliography bug (references appearing as ?).

%opening
% TODO: Contribution of this paper
% Q&A: 'Overlay' or 'Peer-to-Peer'/'P2P'
\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey}
\author{
Vassilis Moustakas$^1$, H\"useyin Akcan$^2$, Mema Roussopoulou$^1$ and Alex Delis$^1$\\
$^1$Department of Informatics and Telecommunications,\\
National and Kapodistrian University of Athens\\
\mbox{\texttt{\{b.moystakas, mema, ad\}@di.uoa.gr}}\\
$^2$Department of Software Engineering,\\
Izmir University of Economics, Izmir, Turkey \\
\mbox{\texttt{huseyin.akcan@ieu.edu.tr}}
}

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% TODO: Contribution of this paper
Peer-to-peer (p2p) systems is a rapidly growing application architecture. The overlay network abstraction on top of a best-effort infrastructure provides a powerful tool to realize ever-wanted exotic features like anonymity, high availability and robustness, load balancing, QoS at the application layer and many more. Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm. One of them is the topology mismatch problem between the overlay and the physical underlying networks, a problem that imposes unnecessary stress to the network resources. This survey investigates on the recent research developments towards the alleviation of the problem, in both unstructured and structures p2p schemes.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

% TODO: Contribution of this survey
% UNDER CONSTRUCTION

% TODO: Structure of this paper
% UNDER CONSTRUCTION!
This survey begins with a quick review of the conception of the overlay network and the available architectures of p2p systems. It continues with the identification of the phenomenon of the topology mismatch between overlay and underlying networks and its impact on the utilization of network resources. Ultimately, it reviews the most important resources found in the literature that attempt to tackle the problem as it proposes a taxonomic scheme, using a set of supertype-subtype relationships based on unique characteristics they have and/or specific goals they target.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OK
\section{Overlay Networks and P2P Systems}

\begin{figure}[h!]
\centering
  \includegraphics[scale=0.5]{img/p2p.pdf}
\caption{An example overlay network.}
\label{figure:overlay}
\end{figure}

An \emph{overlay network} is an abstract, logical interconnection of entities
formed on top of another network. It consists of a set of nodes, the computing
elements, that are being connected by virtual (logical) point-to-point links.
For example, the \emph{Asymmetric Digital Subscriber Line (ADSL)} is, actually,
an overlay network on top of the \emph{Public Switched Telephone Network
(PSTN)}. Figure \ref{figure:overlay} illustrates an example physical ethernet network
and the corresponding overlay network formed by four nodes. 
Nodes $A$ and $B$ are on the same local network, while $C$ and $D$ are in
different networks. The top layer represents the overlay network formed by these
nodes, which is a reflection of the application layer connections among these
nodes. As the topology of the network can change based on the application, in
this example, it is easy to observe that even though nodes $A$ and $B$ are in the same
network, their communication goes through node $C$ on another network,
demonstrating a non-optimal mapping between the physical and the overlay
networks.
An overlay network is generally useful when
abstraction of the network layer from the application layer is important for
applications, particularly when the underlying network structure is subject to
frequent changes.%, which is a common case for ad-hoc networks.

The abstraction of overlay networks has been proposed as a way to implement
efficient, fully distributed, application layer services on top of a best-effort
IP layer forming a widely ranged family of protocols collectively known as
\emph{peer-to-peer} or \emph{p2p} for short. P2p network systems gained
significant attention, in recent years, due to their plethora of unique features
in supporting file sharing among huge numbers of (inter-) networked computers
called \emph{peers}, where each, such, \emph{peer} could act both as a resource
provider and a resource consumer. This changed the traditional
\emph{client-server} model dominating the internet and lead to the introduction
of the \emph{servent}-concept \cite{gnutella}, a portmanteau that blends the
notions of \emph{server} and \emph{client} to denote the twofold role of the
participants in a p2p network. Such serverless systems, proved to be able to
achieve outstanding aggregate resource capacities as participants join the
system
%\footnote{Unfortunatelly, there has been observed mitigation of this
%self-scaling property by the undesirable behaviour of participating users,
%usually answered in popular file-sharing systems; from \emph{free-riding} (i.e.
%in Napster and Gnutella networks \cite{saroiu_measurefileshare_2002,
%adar_gnutellafreeriders_2000, hughes_gnutellafreeride_2005}), to the
%distribution of illegal content and/or other \emph{socio-technical} issues
%\cite{hughes_socp2p_2008}.}, 
without requiring additional expenditure for infrastructure.

As p2p systems evolved, three main architectures have been emerged, namely:
\begin{itemize}
  \item \emph{centralized}
  \item \emph{decentralized unstructured}
  \item \emph{decentralized structured}
\end{itemize}

\emph{Centralized} architectures were the first to recognize that requests for popular content need not be sent to a central server but instead could be handled by the many hosts that already posses the content. \emph{Napster}, for example, maintained a \emph{central index server} based on file lists provided by participating peers. The central index server was queried by the users and it returned pointers to the actual content. Thus, by centralizing search while distributing downloads, Napster achieved a highly functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever''. This centralized search facility, though, renders this scheme not fully distributed and can be proven its ``Achilles' heel'' in terms of being a scalability bottleneck, a central point of failure and vulnerable to malicious acts (e.g. \emph{Denial-of-Service (DoS) attacks}).

The centralized approach has been eventually replaced by architectures that distributed both the search and the download capabilities. In these \emph{decentralized unstructured} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}. The most important properties of such systems are that they support inherent heterogeneity of peers, are highly resilient to peer failures, and incur low maintenance overhead at handling the dynamics of peer participation \cite{stutzbach_churn_2006}. In the literature, they are also known as \emph{broadcast-based} systems, because they use \emph{message flooding} among peers (i.e. Gnutella\cite{gnutella}) or among super-peers (i.e KaZaA\cite{kazaa}) to propagate queries. 
Even though the unstructured approach became highly popular among file sharing applications, they do have a certain disadvantage in locating rare objects due to the unstructured nature and the limitations of the flooding approach. 
%Each (super-) peer receiving a query decreases the incoming query message's lifetime which is measured in \emph{time-to-live (TTL)} hops, back-propagates its results, and  if the lifetime of the incoming message did not expired (TTL did not reach zero), it forwards the message to all its neighbours except from the one it receives it from.
% TODO: maybe enumerate applications of decentralized unstructured p2ps.

Recently, \emph{decentralized structured} schemes have been proposed in order to provide a self-organising infrastructure for large-scale p2p applications \cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002}. They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism. The main advantage of the DHT approach of decentralized structured p2p networks over decentralized unstructured ones, is that they provide a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located even in the case where only a single copy exists in the system. This is $O \left ( log n \right )$ compared to Gnutella, that requires $O \left ( n \right )$ to reliably locate a specific object.
% TODO: maybe enumerate applications of decentralized structured p2ps. (OceanStore, Far)

%{\sethlcolor{yellow}\hl{HA: Give the brief advantages and disadvantages of structured and unstructured
%approaches here!}}

% TODO: enumerate applications of overlay networks and p2p systems regardless of architecture
% UNDER CONSTRUCTION
% Q&A: Need more??
Initial efforts targeted on the implementation of one-to-many addressing schemes
that could replace \emph{IP multicast} in providing higher quality of streaming
media through \emph{quality of service (QoS)} guarantees. IP multicast as well
as other proposals such as \emph{IntServ} and \emph{DiffServ}
\cite{cisco_diffserv_2005} have not seen wide acceptance (yet?), largely because
they require changes to all routers of the network. On the other hand, an
overlay network can be incrementally deployed on end-systems running the overlay
protocol software, without cooperation from the ISPs. Academic research includes
\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.  Overlay
networks were also implemented in order to back the routing of messages to
destinations whose address is not known in advance. This resulted in the
appearance of well known peer-to-peer protocols such as \cite{gnutella} and
\cite{maymounkov_kademlia_2002} widely used for digital content sharing among
their network nodes.  Other special purpose overlay networks include
\cite{anderson_ron_2001} for resilient routing; \cite{subramanian_overqos_2004}
for quality of service guarantees; and \cite{clarke_freenet_2001} for anonymity;
to name just a few.

The main focus of this paper is the topology mismatch problem, therefore the
available approaches in overlay networks are briefly summarized, and interested
readers are pointed to the available surveys on overlay networks
\cite{TheotokisS04,LuaCPSL05}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% start
%TODO: Maybe this section should be reviewed or deleted or go to another section
%Studying the behaviour of the various peer-to-peer schemes \cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003} showed that the ad-hoc network topology of unstructured overlay networks that preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power Law describes the node degree while the Small World describes characteristics of path length and clustering coefficient. The clustering coefficient for a node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of the existing connections between $\upsilon$'s neighbouring nodes to $\gamma \times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring nodes of $\upsilon$. High cluster coefficient means that neighbouring nodes of any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999, saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
%\begin{itemize}
%  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured systems can have high maintenance traffic in delivering messages, updating the mapping, discovering failures and replicating lost data or pointers, making them insufficient on highly volatile networks.
%  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs there is a tight control between the data placement and the topology of the network. For this reason it is hard to efficiently support partially matched queries while Gnutella and other similar systems effortlessly support keyword searches and other complex queries since the mechanism is realized locally, on a node-by-node basis.
%  \item Popular content is located at multiple peers and thus it is more likely for a flooding-based search to return results. DHTs, on the other hand, fit better in the systems which require ability to reliably locate content, even in the extreme case that only a single-copy exists in the network.
%\end{itemize}
% end
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% TODO: how 
%For this reasons, efforts have been placed for optimizing the efficiency of decentralized unstructured peer-to-peer networks. Research mainly focuses on
%\begin{inparaenum}[\itshape i\upshape)]
%  \item reducing unnecessary, redundant communication traffic, and
%  \item exploiting physical locality to reduce communication response.
%\end{inparaenum}
%The goal can be achieved at, both, the application-level network as well as the underlying physical one. In the first case by refining the message relay techniques, while in the second one, by adaptively reconstructing the application network to map as well as possible to the the physical network.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Topology Mismatch Problem}

% TODO: Define the problem
% UNDER CONSTRUCTION
% Q&A is it too late??

%{\sethlcolor{yellow}\hl{ {\cite{hsiao_redblue_2009}} states that:
%Liu {\cite{liu_thancs_2008}}  concluded that more than 70 percent of communication
%paths in an unstructured overlay do not exploit the physical network topology
%(i.e., the Internet topology), resulting in a topology mismatch between the
%overlay and the underlying network.  } }

One of the major issues that defines the efficiency of the overlay network is
the mapping of the physical links and the overlay paths. Ideally, overlay
networks are expected to achieve an optimal mapping among the overlay paths and
the underlying physical links, and avoid the inefficient mapping that is shown
in Figure \ref{figure:overlay}. The problem of constructing an optimal overlay
is referred to as the \emph{topology mismatch problem}, and formally defined as
follows:

\begin{definition}
    Let $V = \{v_1, ..., v_n\}$ be a set of points denoting the network nodes,
    $\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and
    $v_j$, $G=(V,E)$ be a complete distance graph over $V$. The topology
    mismatch problem is to construct a minimal spanning tree,  where 
    node degree is restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
\end{definition}

Constructing degree constrained spanning tree is known to be NP-Hard
\cite{NPBOOK}, as well as the topology mismatch problem
\cite{chawathe_scattercast_2000}. 
 Moreover, the internet is an interesting environment
where end-to-end latencies demonstrate triangle inequality violations (TIVs),
which further complicates the problem.
These delay space TIV is a consequence of the Internet's structure and routing
policies and thus will remain a property of the Internet for the foreseeable
future \cite{zheng_irprtt_2005}. TIVs affect network coordinate
\cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
and makes difficult the construction of delay aware overlays.  A number of
heuristic approaches have been proposed in an effort to finding an algorithm
that performs reasonably and consistently well.

A topology unaware overlay network is able to control the sequence of peers a
message traverses before reaching its destination, but it is completely unaware
of how the actual packets are switched at the underlying infrastructure along
the overlay path. For example, a single logical point-to-point link on the
overlay, most of the time, corresponds to multiple physical links in the
underlying layer. Additionally, a link in the underlying network often serves
the mapping of several overlay paths causing increase of the traffic on the
physical link, which is also called the link's
\emph{stress}\cite{chu_esm_2002}.  Furthermore, the stochastic behaviour of
peers randomly joining and leaving the peer-to-peer network all cause the
overlay to physical mapping to create such unnecessary, redundant maintenance
traffic that impacts the efficiency of the network in terms of average response
time.

%{\sethlcolor{yellow}\hl{
%HA: Motivation to solve topology mismatch problem:
%
%Most of the traffic (find refs!! see Location aware topology matching paper
%introduction for refs [31][29]) in the internet today is generated by P2P
%networks. The inefficient use of the network resources by the P2P protocols,
%especially the problem known as topology mismatch problem, causes unnecessary
%traffic load and a heavy stress
%on the internet infrastructure in general, and effects other users of the
%internet than the P2P nodes.
%}}

\begin{figure}
\centering
  \includegraphics[scale=0.8]{img/phys.pdf}
\caption{Interconnection of nodes in the physical level.}
\label{figure:phys}
\end{figure}

\begin{figure}
\centering
  \includegraphics[scale=0.8]{img/over1.pdf}
  \hspace{3ex}
  \vline
  \hspace{3ex}
  \includegraphics[scale=0.8]{img/over2.pdf}
\caption{HA: Put subfigure, and a new caption here}
\label{figure:over1}
\end{figure}

For example, assume nodes $A$, $B$, $C$ and $D$ are connected through the
physical network shown in Figure~\ref{figure:phys}, where the network costs are
given in milliseconds, and peer $A$ sends a message to peer $D$. If these peers
participate in an overlay network according to one of the setups of
Figures~\ref{figure:over1}~(left) and~\ref{figure:over1}~(right) then users will
yield different performances. In the first overlay shown in Figure
\ref{figure:over1}~(left), the message will traverse the following sequence of
links in the physical layer (marked with their costs): $3 \rightarrow 7
\rightarrow 1 \rightarrow 2 \rightarrow 2 \rightarrow 1 \rightarrow 7
\rightarrow 8 \rightarrow 5 \rightarrow 5 \rightarrow 4$. In the second overlay
shown in Figure ~\ref{figure:over1}~(right), the path will be: $3 \rightarrow 8
\rightarrow 4$. The total cost of the first path is $45 ms$, while the second
only costs a mere $15 ms$. Therefore, we can conclude that the second overlay is
more congruent with the underlying physical network than the first one and thus
more efficient.

% TODO: Argue on the problem
% UNDER CONSTRUCTION
Early incarnations of the overlay protocols, however, did not make use of the optimal
mapping with the physical network topology. Early Gnutella protocol, though were
considered far from scalable \cite{ritter_gnucantscale_2001}, randomly choose
logical neighbours without knowledge of the underlying network, effectively
causing mismatch between physical and application-level topologies.
Additionally, queries may be flooded to multiple paths merging in one peer while
in such case one of the paths would have been enough. 
%Nevertheless, peers may
%forward the same message to each other before they receive the query messages
%from each other.

Similar problems are observed in decentralized structured schemes also.
Typically, node IDs are assigned \emph{randomly}, resulting in excellent load
balancing, scalability and robustness.  Unfortunately, this randomness has a
negative impact on the \emph{routing locality} of the network. This means that
even though the target node can be reached with logarithmic overlay hops, the
distance traveled in the physical underlying network, during the overlay routing
process, can be far from optimal.  The ratio of the physical distance, a query
travels through the overlay to an object, and the minimal distance to that
object (i.e. though IP) is known in the literature as \emph{stretch} or
\emph{Relative Delay Penalty (RDP)} \cite{chu_esm_2000}.

Studies on peer-to-peer traffic reveals the effect of the peer-to-peer traffic
on the overall internet.  Measurements
\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004} on some popular
systems, such as FastTrack, Gnutella and DirectConnect have shown that the
peer-to-peer traffic contributes the largest portion of the overall Internet
traffic.  \cite{matei_mapgnutella_2002} showed that, even given that 95\% of any
two nodes are less than 7 hops away from each other, a flooding-based algorithm,
like the one used by Gnutella, can generate 330TB/month in a 50.000 node
network!  Therefore, the efficiency of the overlay networks has serious
consequences on the overall well-being of the modern internet.

@@

%Caching based protocols are effectively used to reduce traffic costs and response
%times. The caching policy varies depending on the way protocol handles the index and the
%content. Centralized P2P systems
%use central index servers, while local caching systems, such as KazaA, use super peers

@@

%information to detect proximity. The brief description of each category is presented
%below:

@@

%%\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Forwarding} & \textbf{Cache} & \textbf{Overlay optimization} & \textbf{Proximity information} & \textbf{Base protocol} & \textbf{Dynamic update} & \textbf{Runtime} \\

@@

Having understood the nature of the problem that peer-to-peer architectures face,
Section \ref{sec:unstructured} and \ref{sec:structured} discuss the recent academic work that has been conducted
in the field. What follows does not claim
to be a thorough citation of all known protocols that are available out there,
but instead, a careful selection of those that left a distinctive fingerprint
contribution in the efforts of the research community to alleviate the topology
mismatch problem.

%{\sethlcolor{yellow}\hl{
%Create a theoretical section. This is an Approximation alg for Top-mis problem
%
%A Near-Optimal Algorithm Attacking
%the Topology Mismatch Problem
%in Unstructured Peer-to-Peer Networks
%}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Unstructured Decentralized Algorithms}\label{sec:unstructured}

In this section, the unstructured decentralized P2P algorithms are analysed based on their architecture and
base methodology to tackle the topology mismatch problem. The
unstructured decentralized algorithms are visited and categorized based on their
use of the overlay structure, messaging between nodes, and on the techniques
proposed to detect proximity to optimize the topology. 
The base methods that unstructured P2P algorithms use to tackle the topology
mismatch problem can be categorized as Broadcast, Cache, Overlay Optimization, 
and Landmark Based. In this section, these methodologies are discussed in
detail and the algorithms in each category are highlighted.

%{\sethlcolor{yellow}\hl{
%HA: TALK ABOUT GNUTELLA HERE!!!!}}

\subsection{Broadcast Based}

In the naive Gnutella protocol, each received query request is forwarded to all the
neighbours, which clearly generates unnecessary traffic over the network. The
algorithms categorized in the \textit{broadcast based} methodology prefers
re-broadcasting the query messages to a selected subset of their neighbours,
where the selection criteria varies between algorithms.
The selection is made using one or a
combination of various statistical metrics, such as the reliability or the latency
of the link between nodes, etc.
The forwarding based approach enhances search
efficiency but has also the drawbacks of having drastically reduced search scope
due to the selective re-broadcasting policy, and scalability problems on large
P2P networks.
%Expanding the search scope, on the other hand, is no easy task
%because the overhead of forming multicast trees is proportional to the multicast group size.  
%forwarding based schemes do not consider dynamic joining
%and leaving of peers so they do not scale well on dynamic environments.

The forwarding based P2P algorithms and their details are presented below.

\paragraph*{\bf Improving search in peer-to-peer networks}

\cite{yang_improvep2psearch_2002} proposes an easy to implement and practical
solution to the inefficiency problem caused by blind flooding in Gnutella like
P2P file sharing protocols. The paper replaces blind flooding with three
approaches, namely \textit{iterative deepening}, \textit{directed BFS}, and
\textit{local indices}. In \textit{iterative deepening}, the search is performed
on a BFS tree with multiple preset depths. The depth limit is iteratively
increased by the source node for each query based on the quality of the results.
The source node, after examining the results, may issue a new request by
increasing the depth limit, which will trigger the nodes at the last depth
level to resume the search. The iterative approach avoids to start the search
from scratch at each iteration and reduces the load of the nodes on the upper
levels of the tree. The major drawback of \textit{iterative deepening} is the
delay between successive iterations, as the source node needs to examine the
results at each iteration before deciding to quit or resume the query.  The
\textit{directed BFS} tries to avoid this delay by forwarding the query messages
only to a selected set of neighbours, in which the selection criteria varies
from the number of results received previously, distance in terms of hops,
bandwidth, or the query load of the neighbour. In \textit{local indices}, each
node sustains a local data index of the nodes within a radius of $r$ hops of
itself and uses this local index to remotely query the neighbour nodes without
generating additional traffic.  \textit{Local indices} greatly reduces the
aggregate bandwidth usage of the network and improves query efficiency, however,
updating the indices in cases with frequent node joins and leaves introduces a
serious overhead to the system if the radius is kept broad.

\paragraph*{\bf Gia}

\emph{Gia} \cite{chawathe_gia_2003} is an effort to solve the scalability
problem of the unstructured P2P file sharing systems, in particular Gnutella.
The main novelty in the design of Gia is the replacement of blind flooding
approach of Gnutella with random walks \cite{lv_randomwalks_2002}. Although
random walks is a step in the right direction, issuing only a single copy of the
query within the network reduces the search scope, thus affects the success rate
of the query.  In order to overcome this limitation, Gia introduces a
token-based flow control algorithm, which is essentially an intelligent flow
control algorithm that gradually redirects the queries to nodes with more chance
of answering. Gia also acknowledges the heterogenity in peer bandwidth,
processing power, disk speed, etc, of the nodes in P2P networks and uses this
heterogenity when connecting nodes to each other, such that by using a topology
adaptation algorithm, Gia ensures that high capacity nodes have high degrees and
low capacity nodes are within short proximity of high capacity nodes. In order
to prevent overloading of nodes with query requests, Gia uses a token-based flow
control algorithm in which each node announces the number of query requests it
can handle in terms of tokens to its neighbours, so that neighbours only forward
query requests to nodes that they previously received tokens from. Although the
topology adaptation algorithm Gia uses improves the scalability of the network,
it does not help much in solving the topology mismatch problem, since it does
not consider the underlying physical topology.

\paragraph*{\bf Distributed Cycle Minimization Protocol}

The flooding based query approach of Gnutella creates many duplicate packets in
the network. Nodes receive the same query request from various neighbours, which
they generally discard when identified as duplicate. One other reason for
duplicate messages is the cycles in the forwarding paths. These duplicate
packets affect the performance of the P2P network negatively by increasing the
overall resource usage. Moreover, the more active nodes, that have high
capacity, high bandwidth, or contribute more to the network, suffer more from
the overhead caused by the duplicate packets. Therefore, \emph{Distributed Cycle
Minimization Protocol (DCMP)} \cite{zhu_dcmp_2008} is introduced as a method to
remove the cycles and eliminate the duplicate packets, without sacrificing the
connectivity. In DCMP, once a cycle is detected, the most powerful node in the
cycle is elected as the \emph{Gate Peer} and the cycle is cut from a special
link such that the distance of all the nodes within the cycle to the \emph{Gate
Peer} is minimized. The process is managed by using two special message types,
\emph{Information Collection (IC)} and \emph{Cut Message (CM)}. One disadvantage
of DCMP is that since the distance a forwarded message can travel is limited
with the TTL value, which is practically $7$ in Gnutella, the DCMP cannot detect
cycles of length larger than $7$. Even though cycle elimination improves the
network performance, it does not solve the topology mismatch problem.

%\subsection{Cache Based}
%
%Caching based protocols are effectively used to reduce traffic costs and response
%times. The caching policy varies depending on the way protocol handles the index and the
%content. Centralized P2P systems
%use central index servers, while local caching systems, such as KazaA, use super peers
%to cache indices in a distributed way. Content caching is also possible in P2P
%systems, where nodes cache the forwarded content for further retrievals.
%Although caching has the above mentioned advantages,  duplication
%of messages still exist, which limits the scalability of these approaches.
%Therefore, cache based approaches are analyzed in the following categories:
%  \begin{itemize}
%    \item \emph{data index caching},
%    \item \emph{content index caching},
%    \item \emph{centralized}, and
%    \item \emph{local}.
%  \end{itemize}
%

\subsection{Cache Based}
Caching is a widely used methodology to exploit the locality to minimize the
redundant transfer of data, which is successfully used in web servers and file servers.
As peers in a P2P system also work as servers, intuitively it is expected that
P2P file sharing systems can also benefit from caching. However, one of the most
important characteristic properties of P2P systems is that nodes
join and leave frequently, and the lifetime of a query is relatively short
compared to web servers, which makes caching in P2P networks nontrivial. There are
usually two levels of caching possible in file sharing systems, the indices or
pointers to data can be cached, or the data itself can be cached. Caching is
already implemented and successfully used in some commercial P2P systems. One
popular example is KazaA, which uses caching of indices in its super peers.
We present the other protocols that use caching and their details below.

\paragraph*{\bf Replication Strategies in Unstructured Peer-to-Peer Networks}

>>>>

\markboth{}{}{}

\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured
Overlay Networks: A Survey}

\author{
VASSILIS S. MOUSTAKAS\\\emph{Department of Informatics and Telecommunications}\\\emph{National and Kapodistrian University of Athens}\\
H\"USEYIN AKCAN\\\emph{Department of Software Engineering}\\\emph{Izmir University of Economics}\\
MEMA ROUSSOPOULOS\\\emph{Department of Informatics and Telecommunications}\\\emph{National and Kapodistrian University of Athens}\\
AND\\
ALEX DELIS\\\emph{Department of Informatics and Telecommunications}\\\emph{National and Kapodistrian University of Athens}\\
%  Vassilis Moustakas$^1$, H\"useyin Akcan$^2$, Mema Roussopoulou$^1$ and Alex
%Delis$^1$\\
%  $^1$Department of Informatics and Telecommunications,\\
%  National and Kapodistrian University of Athens\\
%  \mbox{\texttt{\{b.moystakas, mema, ad\}@di.uoa.gr}}\\
%  $^2$Department of Software Engineering,\\
%  Izmir University of Economics, Izmir, Turkey \\
%  \mbox{\texttt{huseyin.akcan@ieu.edu.tr}}
}

\begin{abstract}
Peer-to-peer (P2P) systems is a rapidly growing application architecture. The
overlay network abstraction on top of a best-effort infrastructure provides a
powerful tool to realize a plethora of ever-wanted exotic features like
anonymity, high availability and robustness, load balancing, quality-of-service
(QoS) at the application layer and many more. Unfortunately weaknesses of early
effectuations constrained applications from unleashing the full potential of the
paradigm. One of them is the topology mismatch problem between the overlay and
the physical underlying network, a problem that imposes a huge amount of
unnecessary stress to the network resources and, thus, impede applications from
scaling. In this survey we investigate on the recent research developments
towards the alleviation of this problem, in both unstructured and structured
overlay architectures. We visit the various algorithms available in the
lietrature, focusing on the presentation of how each of them tries to tackle the
problem. Additionally, we provide a finer grained categorisation of the
approaches based on the generic type of methodology they use in order to
accomplish this goal.
% TODO: Review this after the comparison sections are fixed.
Finally we discuss the advantages, disadvantages and novelties introduced by the
algotithms and offer pictorial comparisons for them in order to further
strengthen and organise understanding of the information and knowledge cited, in
an attempt to make this survey a one-stop-shop reference for anyone interested
in the topic, from researchers and students, to system architects and
developers.
\end{abstract}

\category{}{}{}

\terms{Topology mismatch, distributed, peer-to-peer, P2P, overlay}

\keywords{Topology mismatch problem, distributed algorithm, peer-to-peer
architecture, overlay network }

\begin{document}

\begin{bottomstuff}
SAY SOMETHING COOL HERE...
\end{bottomstuff}

\maketitle

\input{intro}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% start
%TODO: Maybe this section should be reviewed or deleted or go to another section
%Studying the behaviour of the various peer-to-peer schemes
%\cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003}
%showed that the ad-hoc network topology of unstructured overlay networks that
%preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power
%Law describes the node degree while the Small World describes characteristics
%of path length and clustering coefficient. The clustering coefficient for a
%node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of
%the existing connections between $\upsilon$'s neighbouring nodes to $\gamma
%\times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring
%nodes of $\upsilon$. High cluster coefficient means that neighbouring nodes of
%any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999,
%saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
%\begin{itemize}
%  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured
%systems can have high maintenance traffic in delivering messages, updating the
%mapping, discovering failures and replicating lost data or pointers, making
%them insufficient on highly volatile networks.
%  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs
%there is a tight control between the data placement and the topology of the
%network. For this reason it is hard to efficiently support partially matched
%queries while Gnutella and other similar systems effortlessly support keyword
%searches and other complex queries since the mechanism is realized locally, on
%a node-by-node basis.
%  \item Popular content is located at multiple peers and thus it is more likely
%for a flooding-based search to return results. DHTs, on the other hand, fit
%better in the systems which require ability to reliably locate content, even in
%the extreme case that only a single-copy exists in the network.
%\end{itemize}
% end
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% TODO: how 
%For this reasons, efforts have been placed for optimizing the efficiency of
%decentralized unstructured peer-to-peer networks. Research mainly focuses on
%\begin{inparaenum}[\itshape i\upshape)]
%  \item reducing unnecessary, redundant communication traffic, and
%  \item exploiting physical locality to reduce communication response.
%\end{inparaenum}
%The goal can be achieved at, both, the application-level network as well as the
%underlying physical one. In the first case by refining the message relay
%techniques, while in the second one, by adaptively reconstructing the
%application network to map as well as possible to the the physical network.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{The Topology Mismatch Problem}

%%%%%%%
%%%%%%% TODO: NEEDS REWORDING
%%%%%%% Additional Material
%%%%%%%
%%%%%%%
%%%%%%%
% P4P [16],instead, aimed to provide a general framework for P2P
%applications by developing ISP-application interfaces so that
%application software can use the network layer information,
%such as congestion status and network topology, for better
%performance and resource utilization.

%P4P [37] project attempts to address
%the problem through custom trackers, both for ISPs and P2P
%systems, using an interface based on a primal-dual decomposition
%of an optimization problem. This interface design simplifies
%the realization of traffic-engineering objectives from each parties’
%perspective and ensures the extensibility of the approach. Through
%simulation-based studies and limited experimental deployments,
%these collaborative approaches have been shown to effectively
%reduce network costs while minimally impacting application performance.
%A clear advantage of these proposals is that they
%allow ISPs to incorporate aggregated traffic policies in their tracker
%recommendations (e.g., a particular traffic balance ratio between
%peering providers). However, all of them require deployments
%of oracles for each participating ISP and their effectiveness is
%ultimately predicated on their adoption by P2P applications and a
%trust relationship between P2P users and their ISPs.

@@

%Caching based protocols are effectively used to reduce traffic costs and
%response times. The caching policy varies depending on the way protocol handles
%the index and the content. Centralized P2P systems use central index servers,
% while local caching systems, such as KazaA, use super peers

@@

%information to detect proximity. The brief description of each category is
%presented below:

@@

%%\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Forwarding} &
%\textbf{Cache} & \textbf{Overlay optimization} & \textbf{Proximity information}
%& \textbf{Base protocol} & \textbf{Dynamic update} & \textbf{Runtime} \\

@@

\input{unstructured}

\input{structured}
